---
title: "MMM Optimization"
output: html_document
date: "2024-08-27"
---

## Step 0: Set up Environment

Load Data and Packages
```{r setup}
library(readxl)
library(ggplot2)
myData <- read_excel("C:/Users/nfole/Downloads/sampledata.xlsx")
```



## Function - Optimize Saturation Curve Paramaters
The “x” variable is a vector of length four and represents the four parameters of the ADBUDG function. The temp1 object calculates the predicted Return from the ADBUDG equation.  That vector is then used in the next line when we calculate the squared error for our predicted values, which are stored in the temp2 object.  Lastly, we sum the temp2 object to get a single value for the sum of squared errors, which is stored in the temp3 object.  The temp3 object is the return object for this function.  When we minimize this function, we are essentially minimizing the sum of squared error, which is actually the same thing that is done in linear regression. 

```{r cars}
Ufun<-function(x, Spend, Return) {
  predictedReturn = x[2] + (x[1] - x[2])*((Spend^x[3])/(x[4] + (Spend^x[3])))
  errorSq = (predictedReturn - Return)^2
  sumSqError = sum(errorSq)
  return(sumSqError)
}
```

## Estimate Diminishing Returns
Define Parameters and Optimize
A – The maximum amount of return possible for a campaign given a long term investment.
B – The minimum amount of return possible for a campaign given a long term investment
C – Controls the shape of the curve.
D – Represents initial market share or market saturation effects.

Return is position. So, A and B min values are 0.   C min value is 1.01 because it is based on ROI.

```{r}
startValVec = c(25000,100,1.5,100000)
minValVec = c(0,0,1.01,0)
maxValVec = c(500000, 500000, 2, 10000000)


optim.parms<-nlminb(objective=Ufun,start=startValVec,
                    lower=minValVec,
                    upper=maxValVec,
                    control=list(iter.max=100000,eval.max=2000),
                    Spend = myData$Spend,
                    Return = myData$Return)
optim.parms
```

Fit Diminishing Returns Curve

```{r}
a = optim.parms$par[1]
b = optim.parms$par[2]
c = optim.parms$par[3]
d = optim.parms$par[4]

curveDFx = seq(from=0, to=max(myData$Spend)*2, length.out=10000)
curveDFy = b+(a-b)*((curveDFx^c)/(d+(curveDFx^c)))
curveDF = data.frame(Spend = curveDFx, Return = curveDFy)
```


## Plot Response Curve

```{r}

maxX = 1.05*max(curveDFx, max(myData$Spend))
maxY = 1.05*max(curveDFy, max(myData$Return))

myPlotDataDF = data.frame(Return = myData$Return, Spend = myData$Spend)
optimLineDF = data.frame(Spend = curveDFx, Return = curveDFy)

scatterPlotPlusFit <- ggplot(myPlotDataDF, aes(x = Spend, y = Return)) +
  geom_point(color="black", shape = 16) +
  theme(panel.background = element_rect(fill = 'grey85'),
        panel.grid.major = element_line(colour = "white")) +
  geom_line(data = optimLineDF, aes(x = Spend, y = Return, color = "darkgreen"))  +
  scale_color_manual(labels = "Optimized ADBUDG Fit",values=c('darkgreen')) +
  theme(legend.title=element_blank(), legend.position = "bottom") +
  coord_cartesian(ylim = c(0,maxY), xlim = c(0,maxX)) +
  ggtitle(paste("Email: ", "Data & Model Fit", sep = " "))

scatterPlotPlusFit

```



## Optimization Set up
Create Function to apply diminishing returns to spend and calculate response

```{r}


adbudgReturn = function(a,b,c,d,Spend){
  adbudgReturn = sum(b+(a-b)*((Spend^c)/(d+(Spend^c))))
  return(adbudgReturn)
}

```

Create Variables for optimization
```{r}

oldSpendVec = myData$Spend
oldReturn = adbudgReturn(a,b,c,d,oldSpendVec)
newSpendVec = oldSpendVec

totalSpend = sum(oldSpendVec)
totalReturn = oldReturn


returnGoal = 600000
increment = 1000

```

## Run Optimization
created a function that evaluates the impact of an extra $1000 in each channel.  The money is then given to whichever channel produced the highest incremental return.  It performs this action over and over again in a while loop until the total return has reached the return goal amount.

```{r}



while(totalReturn < returnGoal){
  incReturns = NULL
  for(i in 1:length(oldSpendVec)){
    oldSpendTemp = newSpendVec[i]
    newSpendTemp = newSpendVec[i] + increment
    
    oldReturnTemp = b+(a-b)*((oldSpendTemp^c)/(d+(oldSpendTemp^c)))
    newReturnTemp = b+(a-b)*((newSpendTemp^c)/(d+(newSpendTemp^c)))
    
    incReturns[i] = newReturnTemp - oldReturnTemp
    
  }
  
  winner = which.max(incReturns)
  newSpendVec[winner] = newSpendVec[winner] + increment
  
  totalSpend = totalSpend + increment
  totalReturn = adbudgReturn(a,b,c,d,newSpendVec)

}

```


Get Returns based on Recommended Spend

```{r}

newReturnVec = b+(a-b)*((newSpendVec^c)/(d+(newSpendVec^c)))
myRecommendedData = data.frame(Campaign = myData$Campaign,
                       Channel = myData$Channel,
                       Return = newReturnVec,
                       Spend = newSpendVec)


```



Analyze Recommendations
```{r}

sum(myRecommendedData$Spend) # Recommended Spend
sum(myRecommendedData$Return)  # Estimated Return from Recommended Spend
sum(myRecommendedData$Spend)/sum(myData$Spend) - 1  # % Increase in Spend to get $600K

```


Recommended vs Current Plot

```{r}

#  Graph current spend vs recommended spend  #
compareDF = data.frame(Campaign = rep(myData$Campaign,2), spendType = rep(c("Actual Spend","Recommended Spend"), each=dim(myData)[1]), Spend = c(myData$Spend, myRecommendedData$Spend))

barChart <- ggplot(data=compareDF, aes(x=Campaign, y=Spend, fill=spendType)) +
  geom_bar(stat="identity", color="black", position=position_dodge())+
  scale_fill_manual(values=c('darkred','darkblue'),
                      name = "") +
  scale_y_continuous(name="Spend") +
  theme(axis.text.x = element_text(angle = 45, hjust = .75)) +
  ggtitle("Breakdown of Spend by Campaign")

barChart



```